### 2012-11-12 paper/appendix.tex
Appendix in a file, but not in paper.

\appendix
\section{Relation to Uniform \(k\)-CFA (A Case Against Acceptability)}
\label{sec:accept}

\cite{dvanhorn:nielson-nielson-popl97} \cite{dvanhorn:Neilson:1999}

This machine's allocation strategy mimics the Uniform k-CFA analysis
in Principles of Program Analysis, which is defined in terms of
``$\delta$ contours.''  However, because the machine represetation makes
context explicit via continuations, we can calculate these contours
rather than thread them throught the evaluator.  In other words, we
can use the CESK* machine without modification to obtain Uniform k-CFA
by way of a simple allocation strategy.  (In this way, it's a
simplification of the presentation in JFP.)

NNH uses a coinductive acceptability relation to specify Uniform
k-CFA:

\[
   C,R \models^{ce}_\delta E
\]

The cache and global environment form a finite store-like structure
holding bindings and return values.  The contour environment ce maps
variables to locations in R which contains their bindings, just as the
environment of the CESK* machine does.  The current contour delta is a
string of application labels describing the enclosing context under
which this term is being analyzed (or evaluated).  If you view the
acceptability relation as a big-step evalator, the
$(\widehat C,\widehat\rho)$ component should be seen as a global
store ce is the environment mapping variables to their locations.

Starting form the initial configuration for a program and iterating
the machine transition relation until reaching a fixpoint of reachable
states will \emph{underestimate} the acceptability relation of Uniform
k-CFA.  You can recover acceptability by feeding this store back into
the initial configuration and iterating again.  Repeating this process
until a complete run of the program reaches no new states will be the
least solution that is acceptable.

HOWEVER.  Why should we care about acceptability?  What this
machine computes is safe.  In other words, it computes a more
precise characterization of the run-time behavior of a program.  In
doing is so, it actually saves work (as can be seen above).

An Example:

\begin{alltt}
 (let ((id (\(\lambda\) (x) x)))
   (begin (id 1) (id 2)))
\end{alltt}

Under Uniform 0-CFA, we would have:
\[
   [{\tt x} \mapsto \{{\tt 1}, {\tt 2}\}] \in \widehat\rho
\]

in the least solution to $\models$.  This says that, when run, 'x' is
bound to 1 or 2.

Under the machine semantics using a 0CFA allocation policy, the trace
semantics of the machine show that x is bound to 1, and that at some
later point, x becomes bound to 1 or 2.  Moreover, the machine would
show that (id 1) evaluates to 1 and only 1, while Uniform 0CFA must
give that (id 1) is either 1 or 2 to be acceptable.  We don't see any
value in these kinds of false flows that are due to the global and
timeless aspects of C,R which acceptability requires the heap to be
both finite and unchanging over the course of abstract
interpretation. (Another view of the difference: the machine abstracts
a program's execution as a \emph{finite state machine} that mimics the
machine interpretation of the program; the aceptability relation of
Uniform \(k\)-CFA abstracts a program's execution as a \emph{finite
  map} that mimics the big-step evaluator: from terms to (sets of)
values.)

\subsection{Another problem with acceptability: Temporal ignorance}

The small-step approach to static analysis brings subtle yet important temporal
richness not found in classical analyses for higher-order programs.
%
Classical analyses (ultimately) compute judgments on program terms and
contexts, e.g., at expression $e$, value $x$ may have value $v$.
%
The judgments do not relate the order in which expressions and context may be
evaluated in a program, e.g., a classical analysis has nothing to say with
regard to question like, ``Do we always evaluate $e_1$ before $e_2$?'' or ``Is
it always the case that a file handle is opened, read and then closed in that
order?''

Small-step analyses, by their nature, encode the temporal relationships between
abstract states.
%
It is sensible to make temporal queries of a small-step analysis.
%
Of course, this does not come for free: respecting temporal order imposes an
order in which states and terms may be evaluated \emph{during} the analysis.
%
Classical analyses can (and do) evaluate expressions in any order, or in some
cases, even in parallel~\cite{might:Prabhu:2010:EigenCFA}.
%
Relaxing that restriction on order affords additional optimizations that we
have \emph{not} performed.

We avoid sacrificing order not simply because we are interested in the
questions it allows us to ask, but because considering temporal order actually
improves the precision of the analysis itself.

\section{Pushdown Analysis}

It is straightforward to instantiate a \emph{pushdown} abstraction by
bounding only the variable binding portion of the heap, but using a
unique allocation strategy for continuations.  Such a strategy
abstracts a program's execution as a \emph{pushdown automata}
that mimics the machine interpretation.  This strategy therefore
models the abstract stack in a true stack like fashion and always
properly matches function calls with their return.

Although such analyses can be formulated straightforwardly in the
abstract machine approach, it is not clear all of the techniques of
this paper can be applied to similar effect in the pushdown context.
The main problem is calculation of an analysis can no longer be
computed as the fixed point of the machine transition relation.
Although there are several implementations (CFA2,ICFP'12), they
operate at speeds roughly on par with our starting point: unoptimized
store widened
machines. \cite{dvanhorn:Earl2012Introspective,dvanhorn:Vardoulakis2011CFA2}

\section{Proof: laziness is precision-preserving}
Given widened machine configurations, we can show that lazy
non-determinism is precision-preserving in the cases that application
positions are store-allocated and not. We first show the latter since
it uses less cluttered transition rules. The high level is that the
``fan-out'' of non-determinism is collapsed back in the store (in the
store-allocated applications case) or delayed a single step by
laziness that would have happened in a step of strictness.

Details of lazy machine given as a diff from the strict machine. \\
State space of $lazy-\widehat{CESK}^*_t$ varies just in
$\mathit{Value}$ (storable values do not change):
\begin{align*}
\mval \in \mathit{Value} &::= z \mid b \mid o \mid \saddr{\maddr} \\
\end{align*}

We define a relation between strict and lazy machines that illustrates
the quotient that the lazy states embody.

\newcommand{\vapprx}[3]{#1 \cong_{#2} #3}
\newcommand{\kapprx}[3]{#1 \cong_{#2} #3}
\newcommand{\lapprx}[3]{#1 \approx_{#2} #3}
\newcommand{\apprx}[3]{#1 \sim_{#2} #3}
\newcommand{\capprx}[2]{#1 \sim #2}

\begin{mathpar}
\inferrule{\mval' \in force(\msto, \mval)}
          {\vapprx{\mval}{\msto}{\mval'}} \quad
\inferrule{\ans{\mval} \in cs}{\lapprx{\ans{\mval}}{\msto}{cs}} \quad
\inferrule{\ev{\mexp, \menv, \mkont, \mcntr} \in cs}
          {\lapprx{\ev{\mexp, \menv,\mkont, \mcntr}}{\msto}{cs}} \\
\inferrule{\co{\mkont, \mval'} \in cs \\
           \vapprx{\mval}{\msto}{\mval'}}
          {\lapprx{\co{\mkont, \mval}}{\msto}{cs}} \quad
\inferrule{\ap{\mval_0, \mval_1, \mkont} \in cs \\
           \vapprx{\mval_1}{\msto}{\mval_1'}}
          {\lapprx{\ap{\mval_0, \mval_1', \mkont}}{\msto}{cs}} \\
\inferrule{\ev{\mexp, \menv, \mkont} \in lcs}
          {\apprx{\ev{\mexp, \menv,\mkont}}{\msto}{lcs}} \quad
\inferrule{\ans{\mval} \in lcs \\
           \vapprx{\mval}{\msto}{\mval'}}
          {\apprx{\ans{\mval'}}{\msto}{lcs}} \\
\inferrule{\co{\mkont, \mval} \in lcs \\
           \vapprx{\mval}{\msto}{\mval'}}
          {\apprx{\co{\mkont, \mval'}}{\msto}{lcs}} \quad
\inferrule{\ap{\mval_0, \mval_1, \mkont} \in lcs \\
           \vapprx{\mval_1}{\msto}{\mval_1'}}
          {\apprx{\ap{\mval_0, \mval_1', \mkont}}{\msto}{lcs}} \\
\inferrule{\forall lc \in lcs, \lapprx{lc}{\msto}{cs} \\
           \forall c \in cs, \apprx{c}{\msto}{lcs}}
          {\capprx{(lcs, \msto)}{(cs, \msto)}}
\end{mathpar}

We also need metafunctions for adding and removing a store component
from states. These are straightforwardly defined, so just name them
$wn$ and $nw$ respectively.
%% c2vc(<e^l r k d>,s) = <e^l r s k d>
%% c2vc(<v k d>,s) = <v s k d>
%% vc2c(<e^l r s k d>) = <e^l r k d>
%% vc2c(<v s k d>) = <v k d>

The property we show is that all states stay related through
reduction. The main difficulty is in showing the resulting stores are
in fact equal. We do this by showing each lazy reduction corresponds
to a set of strict reductions, for which the union of their output
stores equal the lazy store. Also for strict equality, every strict
reduction corresponds to a lazy reduction (immediate from the relation).

\begin{lemma}[Partition step]
$\forall lc, lcs', cs, \msto, \msto'.$
 if  $\capprx{(\{lc\},\msto)}{(cs,\msto)}$
 and $\forall lc' \in lcs'. wn(lc,\msto) \machstep wn(lc', \msto')$ then
 $\exists cs'. \forall c \in cs. \exists c' \in cs', \msto^c$
 such that $wn(c,\msto) \machstep wn(c', \msto^c)$
 and       $\msto' = \bigcup\limits_{c \in cs}{\msto^c}$
 and       $\capprx{(lcs',\msto')}{(cs',\msto')}$.
\end{lemma}
\begin{proof}
Let $lc' \in lcs'$ be arbitrary.
By cases on $wn(lc,\msto) \machstep wn(lc', \msto')$:
\begin{itemize}
\item{Case $\ev{\svar{\mvar}, \menv, \mkont, \msto} \machstep \co{\mkont,
    \saddr{\menv(\mvar)}, \msto}$: \\ By definition of $\machstep$,
  $wn(lc,\msto)$ steps strictly to each of $cs' = \{\co{\mval, \mkont} \mid \mval \in \msto(\menv(\mvar))\}$.  By definition of
  $\capprx{}{}$, $cs = \{lc\}$. The stores are the same, and by
  definition of $\lapprx{}{\msto}{}$, $\lapprx{lc'}{\msto}{cs'}$.}
\item{Case $\ev{\slit{\mlit}, \menv, \msto, \mkont, \msto} \machstep
            \co{\mkont, \mlit, \msto}$: \\
      Immeditate.}
\item{Case $\ev{\slam{\mvar}{\mexp}, \menv, \mkont, \msto} \machstep
            \co{\mkont, \clos{\mvar, \mexp, \menv}, \msto}$: \\
      Immediate}
\item{Case $\ev[^\mcntr]{\sapp[^\mlab]{\mexp_0}{\mexp_1}, \menv, \mkont, \msto} \machstep
            \ev[^\mcntr]{\mexp_0, \menv, \msto', \kar[^\mcntr_\mlab]{\mexp_1, \menv, \maddr}, \msto'}$: \\
      Where $\maddr, \msto' = \mathit{push}_\mlab^\mcntr(\msto,\mkont)$ \\
      Immediate.}
\item{Case $\ev[^\mcntr]{\sif[^\mlab]{\mexp_0}{\mexp_1}{\mexp_2}, \menv, \mkont, \msto} \machstep
            \ev[^\mcntr]{\mexp_0, \menv, \msto', \kif[^\mcntr]{\mexp_1, \mexp_2, \menv, \maddr}, \msto'}$: \\
      Where $\maddr, \msto' = \mathit{push}_\mlab^\mcntr(\msto,\mkont)$ \\
      Immediate.}
\item{Case $\co{\kmt, \mval, \msto} \machstep \ans{\msto, \mval'}$: \\
      Where $\mval' \in force(\msto, \mval)$ \\
      By definition of $\capprx{}{}$, $cs = \{\co{\kmt, \mval'} \mid \mval \in force(\msto, \mval)\}$.
      By definition of $\machstep$, $wn(lc,\msto)$ steps strictly to
      each of $cs' = \{\ans{\mval}\}$. The stores are the same.
      By definition of $\lapprx{}{\msto}{}$, $\forall lc' \in lcs'. \lapprx{lc'}{\msto}{cs'}$.
      By definition of $\apprx{}{\msto}{}$, $\forall c' \in cs'. \apprx{c'}{\msto}{lcs'}$.
      Thus $\capprx{(lcs',\msto)}{(cs',\msto)}$.}
\item{Case $\co{\kar[^\mcntr_\mlab]{\mexp, \menv, \maddr}, \mval, \msto} \machstep
            \ev[^\mcntr]{\mexp, \menv, \kfn[^\mcntr_\mlab]{\mval',\maddr}, \msto}$: \\
      Where $v' \in force(\msto, \mval)$. \\
      By definition of $\apprx{}{\msto}{}$,
        $cs = \{\co{\kar[^\mcntr_\mlab]{\mexp, \menv, \maddr}, \mval'} \mid \mval \in force(\msto, \mval)\}$.
      By definition of $\machstep, \lapprx{}{\msto}{}$,
        $cs' = \{\ev[^\mcntr]{\mexp, \menv, \kfn[^\mcntr_\mlab]{\mval',\maddr}}
            \mid \co{\kar[^\mcntr_\mlab]{\mexp, \menv, \maddr}, \mval'} \in cs\}$.
      The stores are the same, and the previous imply $\capprx{(lcs',\msto)}{(cs',\msto)}$.}
\item{Case $\co{\kfn[^\mcntr_\mlab]{\mvalx{u}, \maddr}, \mval, \msto} \machstep
            \ap{\mvalx{u}, \mval, \mkont, \msto}$: \\
      Where $\mkont \in \msto(\maddr)$. \\
      Immediate.}
\item{Case $\co{\kif[^\mcntr]{\mexp_0, \mexp_1, \menv, \maddr}, \mval, \msto} \machstep
            \ev{\mexp_0, \menv, \mkont, \msto}$: \\
      Where $\mkont \in \msto(\maddr)$. \\
      By definition of $\machstep$, $\strue \in force(\msto, \mval)$, thus by definition of
      $\capprx{}{}$, the strict machine takes the same step.}
\item{Case $\co{\kif[^\mcntr]{\mexp_0, \mexp_1, \menv, \maddr}, \mval, \msto} \machstep
            \ev{\mexp_1, \menv, \mkont, \msto}$: \\
      Where $\mkont \in \msto(\maddr)$. \\
      By definition of $\machstep$, $\sfalse \in force(\msto, \mval)$, thus by definition of
      $\capprx{}{}$, the strict machine takes the same step.}
\item{Case $\ap[^\mcntr_\mlab]{\clos{\mvar, \mexp, \menv}, \mval, \mkont,\msto} \machstep
            \ev[^{\mcntr'}]{\mexp, \menv', \mkont, \msto'}$: \\
      Where $\menv', \msto', \mcntr' = \mathit{bind}^\mcntr_\mlab(\menv, \msto, \mvar, \mval)$
      By definition of $\apprx{}{\msto}{}$,
        $cs = \{\ap[^\mcntr_\mlab]{\clos{\mvar, \mexp, \menv}, \mval', \mkont} \mid
                \mval' \in force(\msto, \mval)\}$.
      Thus $cs' = \{\ev[^{\mcntr'}]{\mexp, \menv', \mkont}\}$ (clearly $\lapprx{lc'}{\msto'}{cs'}$)  and
      for all $c \in cs$, $wn(c,\msto) \machstep wn(c',\msto\sqcup[\mvar\mcntr' \mapsto \{c.\mval\}])$.
      Thus the union of these stores is equal to $\msto'$.}
\item{Case $\ap[^\mcntr_\mlab]{\mop, \mval, \mkont,\msto} \machstep
            \co{\mval'', \mkont, \msto}$: \\
      Where $\mval' \in force(\msto, \mval)$
        and $\mval'' \in\interpdelta(\mop,\mval')$. \\
      By definition of $\capprx{}{}$ and $\interpdelta$,
        $cs = \{\ap[^\mcntr_\mlab]{\mop, \mvalx{u}, \mkont} \mid
                \mvalx{u} \in force(\msto, \mval),
                \mval'' \in\interpdelta(\mop,\mvalx{u})\}$.
      Thus the strict machine takes the same step.}
\end{itemize}
\end{proof}

\begin{theorem}[Laziness preserves precision]
$\forall lcs,cs,lcs',\msto,\msto'$ if $\capprx{(lcs, \msto)}{(cs,\msto)}$
  and $(lcs,\msto) \machstep (lcs',\msto')$ then there exists $cs'$
  such that $(cs,\msto) \machstep (cs',\msto')$
\end{theorem}
\begin{proof}
By definition of $\machstep$,
$lcs' = \{lc' \mid \exists \msto^{lc}. wn(lc,\msto) \machstep wn(lc', \msto^{lc}), lc \in lcs\}$
For each $lc \in lcs$ let $\hat{c}^{lc} \subseteq cs$ be the
smallest set such that $\lapprx{lc}{\msto}{\hat{c}^{lc}}$. \\ By the
previous lemma, there exists a $\hat{c}^{lc}{}'$ such that $\capprx{(\{lc' \mid wn(lc,\msto) \machstep
  wn(lc', \msto^{lc})\}, \msto^{lc})}{(\hat{c}',\msto^{lc})}$. Let
$cs' = \bigcup\limits_{lc \in lcs}{\hat{c}^{lc}{}'}$ and $\msto' = \bigcup\limits_{lc \in lcs}{\msto^{lc}}$.
Since each step individually is related, the union is related: $\capprx{lcs'}{\msto'}{cs'}$.
By definition of $\machstep$, $(cs, \msto) \machstep (cs', \msto')$ and $(lcs, \msto) \machstep (lcs', \msto')$.
\end{proof}

%% --Eval
%% <x r s k d> --> <v s k d> where v in s(r(x))  [VAR]
%% <(e0^l0 e1^l1)^{l,lf,la} r s k d> --> <e0^l0 r s\sqcup[ld |-> {k}] ar^{l,lf,la}(e1^l1 r ld) d> [APP]
%% <(lambda x e) r s k d> --> <clos(x e r) s k d> [CLOS]
%% --Continue
%% <v s ar^{l,lf,la}(e^le r a) d> --> <e^le r s\sqcup[lfd |-> {v}] fn^{l,la}(lfd, a)> [CO]
%% --Apply
%% <v s fn^{l,la}(fa, ka) d> --> <e^fl r[x |-> xd'] s'[xd' |-> s'(lad)] k d'> [AP]
%%   where clos(x e^fl r) in s(fa)
%%         k in s(ka)
%%         s' = s\sqcup[lad |-> {v}]
%%         d' = truncate(ld, K)

%% This extra binding corresponds directly to the intermediate
%% bindings introduced by ANF.
%% Matt's ANF analyses likely have all the flavours of our "laziness" technique, but this is how we get our chocolate with our peanut butter as Olin would say.

%% Compare this to lazyCESK^*t:
%% v ::= clos(x e r) | addr(a)
%% Clos ::= clos(x e r)
%% s : a -> P(Clos)
%% others the same

%% force(s, addr(a)) = s(a)
%% force(s, clos(x e r)) = {clos(x e r)}

%% --Eval
%% <x r s k d> --> <addr(r(x)) s k d>
%% others the same
%% --Continue
%% <v s ar^{l,lf,la}(e^le r a) d> --> <e^le r s\sqcup[lfd |-> force(s,v)] fn^{l,la}(lfd, a)>
%% --Apply
%% <v s fn^{l,la}(fa, ka) d> --> <e^fl r[x |-> xd'] s'[xd' |-> s'(lad)] k d'>
%%   where clos(x e^fl r) in s(fa)
%%         k in s(ka)
%%         s' = s\sqcup[lad |-> force(s, v)]
%%         d' = truncate(ld, K)
%% The same globalizing stuff applies.

%% Let us relate the state spaces of the two abstract machines for the lambda calculus in the following way

%% Then we can show that if lazyconf ~ conf and lazyconf ==> lazyconf' then there exists a conf' such that conf ==> conf' and lazyconf' ~ conf'
%% Let lc in lazyconf.cs be arbitrary (call lazyconf.s, s)
%% By cases on lc:
%% <x r k d>:
%% Premises:
%% <x r k d> in conf.cs
%% Thus:
%% Since
%% <x r s k d> -->lazy   <addr(r(x)) s k d>
%% <x r s k d> -->strict <v s k d> where v in s(r(x))
%% By definition of force, v in force(s, addr(r(x)))
%% Thus the first rule of ~_s applies and the second rule of l~_s applies.

%% <(lambda x e) r k d>:
%% Trivial

%% <(e0^l0 e1^l1)^{l,lf,la} r k d>:
%% Premises:
%% <(e0^l0 e1^l1)^{l,lf,la} r k d> in conf.cs
%% Thus:
%% <(e0^l0 e1^l1)^{l,lf,la} r s k d> -->lazy   <e0^l0 r s\sqcup[ld |-> {k}] ar^{l,lf,la}(e1^l1 r ld) d>
%% <(e0^l0 e1^l1)^{l,lf,la} r s k d> -->strict <e0^l0 r s\sqcup[ld |-> {k}] ar^{l,lf,la}(e1^l1 r ld) d>

%% Stores are same in this case.

%% <v ar^{l,lf,la}(e^le r a) d>:
%% Premises:
%% For all v' in force(s, v),
%%   <v' ar^{l,lf,la}(e^le r a) d> in conf.cs
%% Thus:
%% <v ar^{l,lf,la}(e^le r ka) d> -->lazy   <e^le r s\sqcup[lfd |-> force(s, v)] fn^{l,la}(lfd ka) d>
%% for all v' in force(s,v),
%% <v' ar^{l,lf,la}(e^le r ka) d> -->strict <e^le r s\sqcup[lfd |-> {v'}] fn^{l,la}(lfd ka) d>
%% Thus the union of all the strict stores will equal the lazy store.

%% <v fn^{l,la}(fa ka) d>:
%% Premises:
%% For all v' in force(s,v),
%%   <v' fn^{l,la}(fa ka) d> in conf.cs
%% Thus
%% <v fn^{l,la}(fa ka) d> -->lazy <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> force(s,v)]
%% For all v' in force(s,v),
%% <v' fn^{l,la}(fa ka) d> -->strict <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> {v'}]

%% Thus the union of all the strict stores will equal the lazy store.